Surrogate waveform model using Gaussian process regression for an aligned-spin effective one body model of binary neutron star inspirals

### 2018-11-15 paper/paper.tex
Paper : worked on appendix on TF2 phasing.

The explicit expressions for the amplitude and phase of the TaylorF2 waveform are as follows. For the amplitude we use the 1PN correction to the leading order waveform.
\begin{equation}
A_{\rm F2} = \sqrt{\frac{5\pi\eta}{24}} x^{-7/4} \left[ 1 + \left(-\frac{323}{224} + \frac{451\eta}{168} \right)x \right],
\end{equation}
where $x=(\pi M f)^{2/3}$ is the standard PN parameter. This differs slightly from the decomposition used by \texttt{LAL}~\cite{lal}. The phase has the schematic form
\begin{align}
\Phi_{\rm F2} =& -2\pi f t_c + \phi_c + \frac{\pi}{4} - \frac{3}{128\eta}x^{-5/2} \left[1 + \right. \\
                        & \left. {\rm PP}(\eta) + {\rm Spin}(\eta, \chi_1, \chi_2) + {\rm Tidal}(\eta, \Lambda_1, \Lambda_2)  \right].
\end{align}
We use the point particle terms ${\rm PP}(\eta)$ to 3.5PN order in Eq.~(3.18) of Ref.~\cite{BuonannoIyerOchsner2009}, the aligned spin terms ${\rm Spin}(\eta, \chi_1, \chi_2)$ to 3PN order~\cite{BoheMarsatBlanchet2013}, and the tidal terms ${\rm Tidal}(\eta, \Lambda_1, \Lambda_2)$ to 6PN order~\cite{VinesFlanaganHinderer2011}. These terms for the phase are exactly as used in the \texttt{LAL} waveform \texttt{TaylorF2}.

The spin terms are

\red{Because the \texttt{SEOBNRv4T} waveform does not include 3PN spin-spin effects, we remove the 3PN spin-spin term in the \texttt{TaylorF2} waveform as well. The main reason for doing this is because including only common physical effects in both \texttt{SEOBNRv4T} and \texttt{TaylorF2} makes the residual $\Delta\Phi(Mf; \bx)$ smaller and have less variation, making it easier to fit at a given error requirement.}

$SL = X_1^2\chi_1 + X_2^2\chi_2$

$d\Sigma L = (X_1-X_2)(X_2\chi_2-X_1\chi_1)$

$\gamma = \left(\frac{554345}{1134}+\frac{110\eta}{9}\right)SL + \left(\frac{13915}{84} - \frac{10\eta}{3}\right)d\Sigma L$

$SS15 = \frac{188}{3}SL + 25 d\Sigma L$

$SS2 = -10 \Sigma (not same sigma)$

$SS25 = -\gamma -3\gamma\ln(v)$

$SS3 = \frac{\pi}{3}\left[3760(X_1^2\chi_1 + X_2^2\chi_2) + 1490(X_1-X_2)(X_2\chi_2 - X_1\chi_1)\right]$

\begin{equation}
\tilde\Lambda = \frac{16}{13} \frac{(X_1+12X_2)X_1^4\Lambda_1 + (12X_1+X_2)X_2^4\Lambda_2}{(X_1+X_2)^5}
\end{equation}

\begin{align}
\delta\tilde\Lambda =& \frac{(1319X_1^2 - 7996X_1X_2 - 11005X_2^2)X_1^4\Lambda_1}{1319(X_1+X_2)^6} \\
& + \frac{(11005X_1^2 + 7996X_1X_2 - 1319X_2^2)X_2^4\Lambda_2}{1319(X_1+X_2)^6}
\end{align}

$a_5 = -\frac{39}{2}\tilde\Lambda$

$a_6 = -\frac{3115}{64}\tilde\Lambda + \frac{6595}{364}(X_1-X_2)\delta\tilde\Lambda$

>>>>

The explicit expressions for the amplitude and phase of the TaylorF2 waveform are as follows. For the amplitude we use the 1PN correction to the leading order waveform.
\begin{equation}
A_{\rm F2} = \sqrt{\frac{5\pi\eta}{24}} v^{-7/2} \left[ 1 + \left(-\frac{323}{224} + \frac{451\eta}{168} \right)v^{2} \right],
\end{equation}
where $v=(\pi M f)^{1/3}$ is the standard PN parameter. This differs slightly from the decomposition used by \texttt{LAL}~\cite{lal} \red{[explain]}. The phase has the schematic form
\begin{align}
\Phi_{\rm F2} =& -2\pi f t_{0} + \phi_{0} + \frac{\pi}{4} - \frac{3}{128\eta}v^{-5} \left[\Phi_{\rm F2}^{\rm PP}(\eta) \right. \nonumber\\
	& \left. \qquad + \Phi_{\rm F2}^{\rm Spin} (\eta, \chi_1, \chi_2) + \Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)  \right] \,,
\end{align}
where $t_{0}$ and $\phi_{0}$ are constants related to the freedom of choosing the time and phase of coalescence. We separated point particle terms $\Phi_{\rm F2}^{\rm PP}(\eta)$, spin terms $\Phi_{\rm F2}^{\rm Spin}(\eta, \chi_{1}, \chi_{2})$ and tidal terms $\Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)$. We include point particle terms to the highest know order of 3.5PN~\cite{}.

The spin terms can be decomposed into spin-orbit and spin-spin terms (ignoring cubic-in-spin contributions at 3.5PN~\cite{}) as $\Phi_{\rm F2}^{\rm Spin} = \Phi_{\rm F2}^{\rm SO}  + \Phi_{\rm F2}^{\rm SS} $. We include spin-orbit terms up to 3.5PN~\cite{}. For spin-square terms, as the \texttt{SEOBNRv4T} waveform does not include yet 3PN spin-spin effects (see Sec.~\ref{}), we stop at  the leading 2PN in the \texttt{TaylorF2} waveform and ignore the known 3PN terms. The main reason for doing this is because including only common physical effects in both \texttt{SEOBNRv4T} and \texttt{TaylorF2} makes the residual $\Delta\Phi(Mf; \bx)$ smaller and have less variation, making it easier to fit at a given error requirement.

The tidal terms $\Phi_{\rm F2}^{\rm Tidal}(\eta, \Lambda_1, \Lambda_2)$ are known to 6PN order~\cite{VinesFlanaganHinderer2011}. These tidal contributions to the phase are exactly as used in the \texttt{LAL} waveform \texttt{TaylorF2}.

For completeness, we give below explicit expressions of the PN coefficients entering the phasing, using the notations $\delta \equiv (m_{1} - m_{2})/(m_{1} + m_{2})$ for the mass difference, and $\chi_{s,a} = (\chi_{1} \pm \chi_{2})/2$, $\kappa_{s,a} = (\kappa_{1} \pm \kappa_{2})/2$ for the symmetrized and antisymmetrized combinations of $\chi_{A}$, $\kappa_{A}$.

The point particle terms coefficients up to 3.5PN can be found e.g. in Eq.~(3.18) of Ref.~\cite{BuonannoIyerOchsner2009}. The non-zero coefficients read
\begin{align}
	a_{2} &= \frac{55 \nu }{9}+\frac{3715}{756} \nonumber\\
	a_{3} &= -16 \pi \nonumber\\
	a_{4} &= \frac{3085 \nu ^2}{72}+\frac{27145 \nu }{504}+\frac{15293365}{508032} \nonumber\\
	a_{5} &= -\frac{65 \pi  \nu }{9}-\frac{65}{3} \pi  \nu   \ln v +\frac{38645}{252} \pi   \ln v +\frac{38645 \pi }{756} \nonumber\\
	a_{6} &= -\frac{127825 \nu ^3}{1296}+\frac{76055 \nu ^2}{1728}+\frac{2255 \pi ^2 \nu }{12} \nonumber\\ 
	& -\frac{15737765635 \nu }{3048192}-\frac{6848  \ln v }{21}-\frac{640 \pi ^2}{3}-\frac{6848 \gamma_{E} }{21} \nonumber\\ 
	& +\frac{11583231236531}{4694215680}-\frac{13696 \log (2)}{21} \nonumber\\
	a_{7} &= -\frac{74045 \pi  \nu ^2}{756}+\frac{378515 \pi  \nu }{1512}+\frac{77096675 \pi }{254016}
\end{align}

The spin-orbit and spin-spin corrections have the structure
\begin{align}
	\Phi_{\rm F2}^{\rm SO}(\eta, \chi_{1}, \chi_{2}) &= \sum_{k=3}^{7} b_{k} v^{k} + \mathcal{O}(v^{8})\nonumber\\
	\Phi_{\rm F2}^{\rm SS}(\eta, \chi_{1}, \chi_{2}) &= c_{4} v^{4} + c_{6} v^{6} + \mathcal{O}(v^{7}) \,.
\end{align}
\begin{widetext}
The non-zero spin-orbit coefficients are
\begin{align}
	b_{3} &= \frac{113 \delta   \chi_a }{3}+\left(\frac{113}{3}-\frac{76 \nu }{3}\right)  \chi_s \nonumber\\
	b_{5} &=  \chi_s  \left(\frac{340 \nu ^2}{9}+\frac{24260 \nu }{81}+\frac{340}{3} \nu ^2  \ln v +\frac{24260}{27} \nu   \ln v -\frac{732985  \ln v }{756}-\frac{732985}{2268}\right) \nonumber\\
	& \quad + \chi_a  \left(-\frac{140 \delta  \nu }{9}-\frac{732985 \delta }{2268}-\frac{140}{3} \delta  \nu   \ln v -\frac{732985}{756} \delta   \ln v \right)  \nonumber\\
	b_{6} &= \frac{2270 \pi  \delta   \chi_a }{3}+\left(\frac{2270 \pi }{3}-520 \pi  \nu \right)  \chi_s \nonumber\\
	b_{7} &=  \chi_a  \left(-\frac{1985 \delta  \nu ^2}{48}+\frac{26804935 \delta  \nu }{6048}-\frac{25150083775 \delta }{3048192}\right)+\left(\frac{5345 \nu ^3}{36}-\frac{1042165 \nu ^2}{3024}+\frac{10566655595 \nu }{762048}-\frac{25150083775}{3048192}\right)  \chi_s \,,
\end{align}
while the leading-order quadratic-in-spin term reads
\begin{align}
	c_{4} &= \chi_s ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \nu -50  \kappa_s -\frac{195 \nu }{2}-\frac{5}{8}\right) + \chi_a   \chi_s  \left(-100 \delta   \kappa_s -\frac{5 \delta }{4}+200  \kappa_a  \nu -100  \kappa_a \right) \nonumber\\
	& + \chi_a ^2 \left(-50 \delta   \kappa_a +100  \kappa_s  \nu -50  \kappa_s +100 \nu -\frac{5}{8}\right) \,.
\end{align}
As explained above, to be consistent with the fact that this PN information has not yet been incorporated in \texttt{SEOBNRv4T}, we do not include the next-to-leading 3PN term, which for reference is given by
\begin{align}
	c_{6} &=  \chi_s ^2 \left(-\frac{1495 \delta   \kappa_a  \nu }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \nu ^2-\frac{44255  \kappa_s  \nu }{21}+\frac{26015  \kappa_s }{28}+\frac{3415 \nu ^2}{9}+\frac{829705 \nu }{504}-\frac{1344475}{2016}\right) \nonumber\\
	& + \chi_a   \chi_s  \left(-\frac{1495 \delta   \kappa_s  \nu }{3}+\frac{26015 \delta   \kappa_s }{14}+\frac{745 \delta  \nu }{18}-\frac{1344475 \delta }{1008}-480  \kappa_a  \nu ^2-\frac{88510  \kappa_a  \nu }{21}+\frac{26015  \kappa_a }{14}\right) \nonumber\\
	& + \chi_a ^2 \left(-\frac{1495 \delta   \kappa_a  \nu }{6}+\frac{26015 \delta   \kappa_a }{28}-240  \kappa_s  \nu ^2-\frac{44255  \kappa_s  \nu }{21}+\frac{26015  \kappa_s }{28}-240 \nu ^2+\frac{267815 \nu }{252}-\frac{1344475}{2016}\right)\,.
\end{align}
\end{widetext}

Finally, the tidal contributions to the phasing are given by~\cite{VinesFlanaganHinderer2011}
\begin{equation}
	\Phi_{\rm Tidal}^{\rm SS}(\eta, \Lambda_{1}, \Lambda_{2}) = d_{10} v^{10} + d_{12} v^{12} + \mathcal{O}(v^{13})
\end{equation}
Introducing convenient mass-weighted combinations of the tidal parameters, expressed using $X_{1} = m_{1}/(m_{1} + m_{2})$ and $X_{2} = m_{2}/(m_{1} + m_{2})$, \red{[to be checked]}
\begin{align}
	\tilde\Lambda &= \frac{16}{13} \frac{(X_1+12X_2)X_1^4\Lambda_1 + (12X_1+X_2)X_2^4\Lambda_2}{(X_1+X_2)^5} \nonumber\\
	\delta\tilde\Lambda =& \frac{(1319X_1^2 - 7996X_1X_2 - 11005X_2^2)X_1^4\Lambda_1}{1319(X_1+X_2)^6} \\
& + \frac{(11005X_1^2 + 7996X_1X_2 - 1319X_2^2)X_2^4\Lambda_2}{1319(X_1+X_2)^6} \,,
\end{align}
the coefficients are given by \red{[to be checked]}
\begin{align}
	d_{10} &= -\frac{39}{2}\tilde\Lambda \nonumber\\
	d_{12} &= -\frac{3115}{64}\tilde\Lambda + \frac{6595}{364}(X_1-X_2)\delta\tilde\Lambda \,.
\end{align}
